### 2016-02-19 docs/ms_communitySynchrony.Rmd
adds letters to figures; updates conclusion section and discussion with rare species caveats

\textbf{Corresponding Author}: Andrew Tredennick \\  
Address: Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322 USA \\
Phone: +1-970-443-1599 \\
Fax: +1-435-797-3796 \\
Email: atredenn@gmail.com

@@

Interspecific competition has long been considered a leading driver of species asynchrony (also called compensatory dynamics) [@Elton1942; @Tilman1988; @Ives1995]. 
If two species compete strongly, then fluctuating densities in one species leads to compensatory fluctuations in the densities of the other species [@Lehman2000].
All else being equal, interspecific competition generally desynchronizes population abundances but sunchronizes per capita growth rates [@Loreau2008a; @Loreau2013; @Gross2014], except in the special case where competition is weak and population growth rates are high [@Loreau2013].
When both those conditions are met, the presence of competition synchronizes population abundances [@Loreau2013].
Nonetheless, theoretical and empirical work agree that competition should generally lead to compensatory dynamics in population abundances.

Along with environmental fluctuations and interspecific competition, demographic stochasticity can influence synchrony in small populations.
In the absence of all other forces, and in communities composed of species with similar growth rates, abundances, and demogaphic variance, demographic stochasticity leads to independent species fluctuations and always desynchronizes population dynamics [@Loreau2008a; @DeMazancourt2013a].
However, the effect of demographic stochasticity is strong only in small populations, and in theory it cannot oppose or amplify the effects of environmental fluctuations and interspecific competition [@Loreau2008a].

@@

One potential way forward is to use population models fit to empirical data to understand the influence of different drivers on species synchrony.
@Mutshinda2009 used such an approach by fitting a dynamic population model to several community time series of insect and bird abundances.
They used a statistical technique to decompose temporal variation into competition and environmental components, and found that postively correlated environmental responses among species determined community dynamics.
While a major step forward, Mutshinda et al.'s (2009) modeling technique relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
Thus, despite almost a decade of research aimed toward understanding species synchrony, we still lack direct tests that link environmental fluctuations, interspecific interactions, and demographic stochasticity to species synchrony in natural communities.

@@

Discussion
----------
Temporal asynchrony among interacting populations can stabilize ecosystem-level function and is influenced by inter and intraspecific competition, demographic stochasticity, and environmental fluctuations [@Loreau2013].
We used multi-species population models applied to five semiarid plant communities to determine the relative influence of the drivers of community synchrony.
Our study produced three main findings that were consistent across five plant communities: (1) interspecific interactions were weak and therefore had little impact on community synchrony; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; and (3) environmental fluctuations increased community synchrony since species-specific responses to the environment were positively, but not completely, correlated.
Thus, our results are in agreement with theoretical expectations: when species interactions and demographic stochasticity are weak, community synchrony approximately equals the synchrony of species' responses to the environment.

@@

A challenge to the generality of our results is that a recent analysis of several biodiversity-ecosystem function experiments came to a completely opposing conclusion: competition drives species synchrony in grasslands [@Gross2014].
@Gross2014 found that temporal asynchrony among plant species in four experiments was driven by interspecific competition that counteracted the synchronizing effect of environmental fluctuations.
This is surprising given the extremely weak species interactions we estimated.
But the apparent inconsistency between our results and those of @Gross2014 disappears when we consider the differences between our studies.
Gross et al.'s results are based on experimentally-assembled communities that have coexisted for, at most, 11 years.
In contrast, the natural communities we studied consist of species that have coexisted for decades and have virtually zero niche overlap [@Chu2015].
In theory, species interactions should weaken through time as community assembly progesses [@Kokkoris1999], meaning that in biodiversity-ecosystem function experiments community assembly processes that can result in competitive exclusion may yet to have been realized.
Therefore, such communities may be highly influenced by transient, but strong, interspecific competition that can mask the synchronizing effects of the environment.
This also leads to the prediction that community synchrony is driven by competition in recently assembled communities, while synchrony is driven by environmental fluctuations in communities that have undergone extensive species sorting.
In other words, in the absence of disturbance and invasion, the dominant driver of community synchrony should shift from competition to environmental fluctuations through time.
One way to test this prediction is to continue collecting data from biodiversity-ecoystem function experiments and conduct the analysis of @Gross2014 every few years.

Our analysis also contributes to mounting evidence on the importance of functional diversity for maintaining temporal stability via species asynchrony [@Reich2012; @Polley2013].
Functional traits define how a species will respond to its environment.
Therefore, greater functional differences among species should lead to lower correlations in species' responses to the environment.
We found evidence for the negative effect of functional diversity on community synchrony: community synchrony was lowest Idaho (Table 2, Fig. 2); the only community that contained two life forms (one shrub, three perennial grasses).
In fact, when we ran simulations with the shrub removed, synchrony of the Idaho community increased substantially and was similar to synchrony in the Kansas community, which also contained three species.
Functional diversity is especially important in communities with weak interspecific interactions where synchrony depends soley on species' environmental responses.
Adding species to such communities will always decrease synchrony, and increase temporal stability, so long as the synchrony of species' environmental response is less than 1 [@Loreau2008a; @Loreau2013].
But functionally diverse suites of species will have the lowest synchrony of environmental responses, and thus lowest community synchrony and highest stability.

Conclusion
----------
Species-specific responses to temporally fluctuating environmental conditions is an important mechanism underlying asynchronous population dynamics and, in turn, ecosystem stability [@Loreau2013].
When we removed environmental variation, we found that synchrony decreased (Fig. 3).
A tempting conclusion is that our study confirms that compensatory dynamics are rare in natural communities, and that ecologically-similar species will exhibit synchronous dynamics [e.g., @Houlahan2007].
Such a conclusion is only half correct.
The perennial grasses we studied do have similar responses to the environment (Table X1-X10), which will tend to synchronize dynamics.
However, if community-wide species synchrony is less than 1, some degree of compensatory dynamics must be present [@Loreau2008a].
Since observed and simulated synchrony is less than 1 in all our focal communities (Fig. 1), we conclude that compensatory dynamics are common in natural grasslands.
But, in agreement with other studies, we find that environmental responses are primarily responsible for the degree of synchrony among ecologically-similar species.

We found strong support for Loreau and de Mazancourt's (2008) prediction that community-wide synchrony of per capita growth rates should approximately equal the synchrony of species' environmental responses when demographic stochasticity and species interactions are weak, since removing environmental fluctuations was the only simulate that affected synchrony.
We have reason to believe that our results are quite general since many studies report large niche differences, and thus weak interactions, in plant communities.
However, our results are in contrast to a recent analysis of several biodiversity-ecosystem function experiments showing that competition drives community synchrony [@Gross2014].
The difference between the Gross et al. (2014) study and our own may be due to community assembly processes: recently assembled communities, such as experimental plots, may exhibit strong species interactions that will weaken over time [@Kokkoris1999].
Viewing community synchrony through the lens of community assembly processes may provide a way to reconcile inconsistent observations on the biotic and abiotic forces that jointly drive community dynamics.

>>>>

\textbf{Corresponding Author}: \\
\hspace{2em}Andrew Tredennick \\  
\hspace{2em}Department of Wildland Resources and the Ecology Center\\
\hspace{2em}Utah State University \\
\hspace{2em}5230 Old Main Hill \\
\hspace{2em}Logan, Utah 84322 USA \\
\hspace{2em}Phone: +1-970-443-1599 \\
\hspace{2em}Fax: +1-435-797-3796 \\
\hspace{2em}Email: atredenn@gmail.com

@@

Interspecific competition has long been considered a leading driver of species asynchrony (also called compensatory dynamics) [@Tilman1988; @Ives1995]. 
If two species compete strongly, then fluctuating densities in one species leads to compensatory fluctuations in the densities of the other species [@Lehman2000].
All else being equal, interspecific competition generally desynchronizes population abundances but sunchronizes per capita growth rates [@Loreau2008a; @Loreau2013; @Gross2014], except in the special case where competition is weak and population growth rates are high [@Loreau2013].
When both those conditions are met, the presence of competition synchronizes population abundances [@Loreau2013].
Nonetheless, theoretical and empirical work agree that competition should generally lead to compensatory dynamics in population abundances.
However, competition-induced compensatory dynamics may or may not increase ecosystem stability, depending on how much competition destabilizes population dynamics [@Loreau2013].

Along with environmental fluctuations and interspecific competition, demographic stochasticity can influence synchrony.
In the absence of all other forces, and in communities composed of species with similar growth rates, abundances, and demogaphic variance, demographic stochasticity leads to independent species fluctuations and always desynchronizes population dynamics [@Loreau2008a; @DeMazancourt2013a].
However, the effect of demographic stochasticity is strong only in small populations, and in theory it cannot oppose or amplify the effects of environmental fluctuations and interspecific competition [@Loreau2008a].

@@

One potential way forward is to use population models fit to empirical data to understand the influence of different drivers on species synchrony.
@Mutshinda2009 used such an approach by fitting a dynamic population model to several community time series of insect and bird abundances.
They used a statistical technique to decompose temporal variation into competition and environmental components, and found that postively correlated environmental responses among species determined community dynamics.
While a major step forward, Mutshinda et al.'s (2009) modeling technique relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
Despite almost a decade of research aimed toward understanding species synchrony, we still lack direct tests that link environmental fluctuations, interspecific interactions, and demographic stochasticity to species synchrony in natural communities.

@@

Discussion
----------
Temporal asynchrony among interacting populations can stabilize ecosystem-level function and is influenced by inter and intraspecific competition, demographic stochasticity, and environmental fluctuations [@Loreau2008a].
We used multi-species population models applied to five semiarid plant communities to determine the relative influence of the drivers of community synchrony.
Our study produced three main findings that were consistent across five plant communities: (1) interspecific interactions were weak and therefore had little impact on community synchrony; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; and (3) environmental fluctuations increased community synchrony relative to simulations in constant environments because species-specific responses to the environment were positively, but not completely, correlated.
Thus, our results are in agreement with theoretical expectations: when species interactions and demographic stochasticity are weak, community synchrony approximately equals the synchrony of species' responses to the environment.

@@

One challenge to the generality of our results is that we were only able to model common, co-occuring species [see @Chu2015].
Rare species, which our analysis does not include, could be kept at low densities by competitive suppression due to strong interspecific interactions.
If we had been able to model rare species, we may have found a larger effect of competition on community synchrony.
However, rare species may occupy rare niches that do not overlap with other species, in which case synchrony would remain predominantly driven by environmental responses.
Rare species could also be limited by species-specific natural enemies whose effects are density-independent [@Mordecai2011].
Under such conditions, synchrony will depend on the degree of host-specificity and the pathogen's response to environmental conditions.
Niether our model nor current theory explicitly considers the effects of pathogens on species synchrony, and this highlights the need for theoretical work on the interaction between mechanisms of species coexistence and mechanisms of species synchrony [@Loreau2010].

Another challenge to the generality of our results is a recent analysis of several biodiversity-ecosystem function experiments that came to a completely opposing conclusion: competition drives species synchrony in grasslands [@Gross2014].
@Gross2014 found that temporal asynchrony among plant species in four experiments was driven by interspecific competition that counteracted the synchronizing effect of environmental fluctuations.
This is surprising given the extremely weak species interactions we estimated.
But the apparent inconsistency between our results and those of @Gross2014 disappears when we consider the differences between our studies.
Gross et al.'s results are based on experimentally-assembled communities that have coexisted for, at most, 11 years.
In contrast, the natural communities we studied consist of species that have coexisted for decades and have virtually zero niche overlap [@Chu2015].
In theory, species interactions should weaken through time as community assembly progesses [@Kokkoris1999], meaning that in biodiversity-ecosystem function experiments community assembly processes that can result in competitive exclusion may yet to have been realized.
Therefore, such communities may be highly influenced by transient, but strong, interspecific competition that can mask the synchronizing effects of the environment.
This also leads to the prediction that community synchrony is driven by competition in recently assembled communities, while synchrony is driven by environmental fluctuations in communities that have undergone extensive species sorting.
In other words, in the absence of disturbance and invasion, the dominant driver of community synchrony should shift from competition to environmental fluctuations through time.
One way to test this prediction is to continue collecting data from biodiversity-ecoystem function experiments and conduct the analysis of @Gross2014 every few years.

Our analysis contributes to mounting evidence on the importance of functional diversity for maintaining temporal stability via species asynchrony [@Reich2012; @Polley2013].
Functional traits define how a species will respond to its environment.
Therefore, greater functional differences among species should lead to lower correlations in species' responses to the environment.
We found evidence for the negative effect of functional diversity on community synchrony: community synchrony was lowest Idaho (Table 2, Fig. 2); the only community that contained two life forms (one shrub, three perennial grasses).
In fact, when we ran simulations with the shrub removed, synchrony of the Idaho community increased substantially and was similar to synchrony in the Kansas community, which also contained three species.
Functional diversity is especially important in communities with weak interspecific interactions where synchrony depends soley on species' environmental responses.
Adding species to such communities will always decrease synchrony, and increase temporal stability, so long as the synchrony of species' environmental response is less than 1 [@Loreau2008a; @Loreau2013].
But functionally diverse suites of species will have the lowest synchrony of environmental responses, and thus lowest community synchrony and highest stability.

Conclusions
-----------
Species-specific responses to temporally fluctuating environmental conditions is an important mechanism underlying asynchronous population dynamics and, in turn, ecosystem stability [@Loreau2013].
When we removed environmental variation, we found that synchrony decreased (Fig. 3).
A tempting conclusion is that our study confirms that compensatory dynamics are rare in natural communities, and that ecologically-similar species will exhibit synchronous dynamics [e.g., @Houlahan2007].
Such a conclusion is only half correct.
The perennial grasses we studied do have similar responses to the environment (Table X1-X10), which will tend to synchronize dynamics.
However, if community-wide species synchrony is less than 1, some degree of compensatory dynamics must be present [@Loreau2008a].
Since observed and simulated synchrony is less than 1 in all our focal communities (Fig. 1), we conclude that compensatory dynamics are common in natural grasslands.
But, in agreement with other studies [@Rocha2011; @Vasseur2014], we find that environmental responses are primarily responsible for the degree of synchrony among ecologically-similar species.

We found strong support for Loreau and de Mazancourt's (2008) prediction that community-wide synchrony of per capita growth rates should approximately equal the synchrony of species' environmental responses when demographic stochasticity and species interactions are weak, since removing environmental fluctuations was the only simulation that affected synchrony.
We have reason to believe that our results are quite general since many studies report large niche differences, and thus weak interactions, in plant communities.
However, our results are in contrast to a recent analysis of several biodiversity-ecosystem function experiments showing that competition drives community synchrony [@Gross2014].
The difference between the Gross et al. (2014) study and our own may be due to community assembly processes: recently assembled communities, such as experimental plots, may exhibit strong species interactions that will weaken over time [@Kokkoris1999].
Viewing community synchrony through the lens of community assembly processes may provide a way to reconcile inconsistent observations on the biotic and abiotic forces that jointly drive community dynamics.
